
# Reproducable Research Peer Assesssment 1

## Loading the data

Load the data

```{r}
data <- read.csv("activity.csv")
```

Process/transform the data into a format suitable for analysis

```{r, echo=TRUE}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN = sum, na.rm = TRUE)

```

## What is mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day

```{r, echo=TRUE}
qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")
```

Calculate and report the mean and median total number of steps taken per day

```{r, echo=TRUE}
mean(total.steps)
```

```{r, echo=TRUE}
median(total.steps)
```



## What is the average daily activity pattern?

Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
library(ggplot2)
averages <- aggregate(x = list(steps = data$steps), by = list(interval = data$interval), FUN = mean, na.rm = TRUE)
ggplot(data = averages, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute interval") + ylab("average number of steps taken")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r, echo=TRUE}
averages[which.max(averages$steps), ]
```



## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r, echo=TRUE}
missing <- is.na(data$steps)
# How many missing
table(missing)
```



Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

All of the missing values are filled in with mean value for that 5-minute interval.


```{r, echo=TRUE}
# Replace each missing value with the mean value of its 5-minute interval
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps)) 
        filled <- c(steps) else filled <- (averages[averages$interval == interval, "steps"])
    return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```



Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r, echo=TRUE}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN = sum)
qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")
```

```{r, echo=TRUE}
mean(total.steps)
```

```{r, echo=TRUE}
median(total.steps)
```


Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


: The mean value is the same as the value before imputing missing data because we put the mean value for that particular 5-min interval. The median value shows a little difference : but it depends on where the missing values are.

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) 
        return("weekday") else if (day %in% c("Saturday", "Sunday")) 
        return("weekend") else stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN = weekday.or.weekend)
```

Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:

```{r, echo=TRUE}
averages <- aggregate(steps ~ interval + day, data = filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + 
    xlab("5-minute interval") + ylab("Number of steps")
```

